Enhanced Material Point Method to face dynamic problems:\thanks{Funding: This study was funded by Agustín de Betancourt Foundation (grant number 262390106114).}

### 2020-03-27 _region_.tex
Versión II día 27-3-2020

\message{ !name(main.tex)}%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

\usepackage[lofdepth,lotdepth]{subfig}
% \usepackage{graphicx}
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

% Tikz
\usepackage{stanli}
\usepackage[ugly]{units}

%\usepackage{latexsym}
% etc.
%

@@

\title{Suitability of the local-maximum entropy approximation under
  the Material Point Method framework for dynamic problems. \thanks{Funding: This
    study was funded by Agustín de Betancourt Foundation (grant number
    262390106114).}
}
% \subtitle{}

\titlerunning{Suitability of the LME approximation under
  the MPM framework for dynamic problems.}        % if too long for running head

@@

\maketitle
\begin{abstract}
  This document is devoted to describe the suitability of the Local
  \textit{maximum-entropy} (LME) meshfree approximation technique under the framework of the
  Material Point Method for dynamic problems. 
  \keywords{LME \and MPM \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Since the proposal of the Material Point Method (\textbf{MPM}) by Sulsky {\it
  et al.} (1994)\cite{Sulsky1994}, as a generalization of
a particle-in-cell method \cite{Harlow1956} and a fluid implicit
particle (FLIP) method \cite{Brackbill1986}, MPM has been used to
simulate dynamic problems. However, in the simulations made with the
originalMPM, there are numerical noises \cite{Tran2019e} when it is faced to solve
classic dynamic benchmarks like the one proposed by Dyka \& Ingel
(1995)\cite{Dyka1995}.

The aim of this paper is to mitigate this spurious oscillations by the
employ of the maximum-entropy (or local \textit{max-ent}) shape
function under the MPM framework. First introduced by
\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite elements (\textbf{FE}) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness of the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative, satisfy the partition of unity
property, and provide an exact approximation for affine functions \cite{Arroyo2006}.

This approximation scheme has been proof to have a good performance
under the dynamic regime by other researchers like Navas {\it et al.}
(2018)\cite{Navas2018a} and Li {\it et al.} (2012)\cite{Li2012} for
Optimal Transportation Meshfree (\textbf{OTM}) method.

a comparison between both methods employing \textit{max-ent}  shape
function is performed in this paper

as well a parametric study for the $\gamma$ parameter

finally, conclusions and future research topics are exposed

\section{Notation}
\label{sec:Notation}
In what follows, we will adopt the following convention:
\begin{table}[h]
  \centering
  \caption{Physical variables involved in the problem}
  \label{tab:notation-table}
  \begin{tabular}{c c}
    $\rho$ & Density field \\
    $\vect{a}$ & Acceleration field \\
    $\vect{v}$ & Velocity field \\
    $\vect{u}$ & Displacement field \\
    $\tens{\sigma}$ & Cauchy stress tensor \\
    $\tens{\varepsilon}$ & Cauchy strain tensor\\
    $\tens{D}$ & Constitutive tensor \\
  \end{tabular}
\end{table}

The subscript $p$ is used to define a particle variable. While the subscript $I$ is reserved in this
notation for denoting nodal variables.

\section{Derivation of the MPM}
\label{sec:derivation-mpm}
In this section, we provide an overview of the standard explicit MPM
algorithm as presented by \cite{Sulsky1994}. The method has three main
steps: a variational recovery process to project particle data to the
nodes of the grid, a Eulerian step where balance of momentum is
compute by solving internal and external forces in the nodes, and
finally the advection of the Lagrangian particles. We will first give
a introduction to the governing equations and then follow by an
algorithmic description of the process.

\subsection{Governing equations }
\label{sec:governing-equations}
In the MPM approach a continuum is considered. The continuum behaviour
can be described by a set of governing equations which are

the balance of momentum equation
\begin{equation}
  \label{eq:Balance-momentum}
\rho \vect{a} = \Div{\tens{\sigma}} + \rho \vect{b},
\end{equation}
where $\Div{\qed}$ denotes the divergence operator. To ensure that the
velocity field is compatible with the strain field is necessary that
them satisfy the compatibility equation,
\begin{equation}
  \label{eq:Compatibility-equation}
  \dot{\tens{\varepsilon}} = \GradS{\vect{v}} = \frac{1}{2}\left(
    \Grad{\vect{v}} +  \GradT{\vect{v}} \right),
\end{equation}
where $\GradS{\qed}$ denotes the symmetric part of the gradient
operation. Thorough the constitutive equation, the rate of the Cauchy stress
tensor is defined to be
\begin{equation}
  \label{eq:Constitutive-equation}
\dot{\tens{\sigma}} = \tens{D} \colon \dot{\tens{\varepsilon}},
\end{equation}
where $\tens{D}$ is the classical elasticity tensor. Finally we
introduce the total derivative expression over the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
\frac{D \rho}{D t} = \dot{\rho} + \rho \Grad{\vect{v}}.
\end{equation}
Where if we impose $\frac{D \rho}{D t}$ to be zero, the conservation
of mass is guarantee.

\subsection{Variational formulation}
\label{sec:variational-formulation}
To write the variational statement of the problem, let us define a
virtual velocity field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{v}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{v}^{\psi} \in
  \mathcal{H}^1 \mid \vect{v}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{v}^{\psi}} < \infty\ \quad\text{and}\quad \Integral{3}{\Grad{\vect{v}^{\psi}}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary value problem of elasticity is the function $\vect{v} \in
\mathcal{H}^1_0$ such that, for $\vect{v}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \Deriv{\vec{v}}{t}\ - \vec{b} \right) \cdot \vec{v}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{v}^{\psi}} - \Integral{3}{\tens{\sigma} \cdot
    \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) is the balance of momentum formulated in
its weak form.

\subsection{Galerkin procedure}
\label{sec:Galerkin-procedure}
In order to arrive to a finite set of equations, the continuum domain
$\Omega$ is discretized with a finite sum of material points (in the
following particles), each one represent a part of the discretized domain $\Omega_p$ with $p = 1,2\ldots ,N_p$
where $N_p$ is the number of particles. The material point $\vec{x}_p$ is
defined at the centroid of each $\Omega_p$. Further, a computational grid
is formed of a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, where $N_n$ is the number of grid nodes.

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Discretization_MPM}
  }
  \caption{MPM discretization.}
  \label{fig:MPM-discretization}
\end{figure}

Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,  $\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$. The problem is then integrated using a discrete number of time
steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps.  So employing the definition of the
material integral, individual terms of \eqref{eq:BalanceMomentum_wf}
are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a}\ \vect{v}^{\psi}} =& \sum^{N_p}_{p}
   \dot{\vec{v}}_{p}\ \vect{v}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \Grad{\vec{\psi}}} = \sum^{N_p}_{p}
   \tens{\sigma}_{p}\ \Grad{\tens{\psi}_{p}}\ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b}\ \vect{v}^{\psi} } = \sum^{N_p}_{p}
  \vec{b}_{p}\ \vect{v}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{v}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s\ \vect{v}^{\psi}} = \sum^{N_p}_{p} \vec{t}^s_{p}\ \vect{v}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
and approximating the virtual displacement of the particle $p$ as
$\vec{\psi}_{p} = N_I(\vec{x}_p)\cdot \vect{1} = N_{Ip}$. We reach to
the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} =
  \sum^{Np}_{p} N_{Ip} m_p N_{Jp},
\end{equation}
and internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \sum^{N_p}_{p} \Grad{N_{Ip}}
  \tens{\sigma}_{p} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = \sum^{N_p}_{p} N_{Ip} m_p \vec{b}_{p} + \sum^{N_p}_{p}
  N_{Ip} m_p h^{-1} \vec{t}^s_{p}
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2}\left[\Grad{N_{Ip}}\ \vec{v}_{I} + \Grad{N_{Ip}}^T\
    \vec{v}_{I}\right].
\end{equation}
Later, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

To improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_I$. Therefore the nodal balance of momentum equation \eqref{eq:particle_balance_forces3} can be simplified to
\begin{equation}
  \label{eq:SimplyNodalMomentum}
  \dot{\vec{p}}_{I} = \tens{m}_I\ \vec{a}_{I} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext}.
\end{equation}
To solve the equation \eqref{eq:SimplyNodalMomentum}, a second order
temporal integration scheme is required. The next step is to advect
the particles solving,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = \sum^{N_n}_{I}\vec{a}_{I}\ N_{Ip},\quad and\quad \dot{\vec{x}}_{p} = \sum^{N_n}_{I}\vec{v}_{I}\ N_{Ip}  
\end{equation}
To solve this equations, the forward Euler algorithm is the common
choose for this in the MPM literature. In next lines, a complete
description of the standard MPM algorithm will be described.

>>>>

\message{ !name(main.tex)}%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments

gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
% flush right qed marks, e.g. at end of proof
\smartqed

\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{subfigure} % subfiguras
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

% Tikz
\usepackage{stanli}
\usepackage[ugly]{units}
% \usetikzlibrary{decorations}
% \usetikzlibrary{arrows}
\usetikzlibrary{plotmarks}

\usetikzlibrary{%
    decorations.pathreplacing,%
    decorations.pathmorphing%
}

%\usepackage{latexsym}
% etc.

%

@@

% \title{Novel improvements in the Material Point Method to face
%   dynamic problems: 
%   Local-maximum entropy approximation and an explicit predictor-corrector scheme. \thanks{Funding: This
%     study was funded by Agustín de Betancourt Foundation (grant number
%     262390106114).}
% }

\title{Enhanced Material Point Method to face
  dynamic problems:\thanks{Funding: This
    study was funded by Agustín de Betancourt Foundation (grant number
    262390106114).}
}
\subtitle{Local-maximum entropy approximation and explicit
  predictor-corrector scheme}

\titlerunning{Novel improvements in the Material Point Method to face
  dynamic problems.} % if too long for running head

@@

\maketitle
\begin{abstract}
  Material Point Method (MPM) has arisen in the recent years as an
  alternative to Finite Element Method (FEM) under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard MPM time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper, an explicit predictor-corrector time integration
  scheme has been proposed. Its superior performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in MPM 
  occurs when to material points cross computational grid
  boundaries and motivated due to the lack of smoothness of the interpolation
  functions. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) a robust substitute for the wide range of shape
  function in MPM. Local \textit{max-ent} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying a regularization parameter they
  are able to behave finite element like or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single shape function family. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
  \keywords{LME \and MPM \and Explicit predictor-corrector \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Since the proposal of MPM by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the Fluid
Implicit Particle (FLIP) method~\cite{Brackbill1986}. It popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the generalized interpolation material point method
(GIMP) Bardenhagen \& Kober (2004)~\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost.  Furthermore, as it is  a voxel based
discretization technique, it is prone to suffer voxel domains overlap
or gaps when the material point mesh becomes irregular, which can
introduce severe inaccuracies. This is similar to the difficulty
encountered by the finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)~\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive. In recent years the employ of
spline-lines has gain popularity with the introduction of the B-Spline
MPM proposed by Roel Tielen {\it et al.} (2017)~\cite{TIELEN2017265},
this technique allows the employ of unstructured set of notes and
particles. More recently, approximants derived from minimization has
been introduced in to MPM framework with the Conservative Taylor
Least Squares (CTLS) reconstruction proposed by Wobbes {\it et al.}
(2018)~\cite{E_Wobbes_2018}, unfortunately when particles are spread in
a challenging way, the quality of the CTLS approximation decrease
locally.

This document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) as a robust substitute for the wide range of shape
  function in MPM. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite element method (FEM) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions~\cite{Arroyo2006}. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)~\cite{Navas2018a} and Li {\it et al.} (2012)~\cite{Li2012} for
Optimal Transportation Meshfree (OTM) method. And more
recently under MPM framework by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020} but without exploring the benefits of
the regularization parameter $\beta$.

This techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using
techniques~\cite{Tran2019e}. These numerical inaccuracies also known
as wiggles are due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises is to add a
nonphysical damping source to the equilibrium equations, this
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing the accuracy of the time integration
scheme. One of the most popular is the implicit GIMP (iGIMP)
Charlton {\it et al.} (2017)~\cite{Charlton2017}, more recently Tran \&
Solowski (2019)~\cite{Tran2019e} proposed a generalised-$\alpha$ scheme
for MPM with promising results but at the expense of increasing
the computational effort. In this paper a less time consuming and high
efficient explicit predictor corrector integration method has been
proposed. It consists in an accommodation of the Newmark
predictor-corrector or central difference explicit (CD). We have
choose this method among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because it simplicity and it good
performance dealing with solids dynamics problems under a mesh-free
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employ of a
suitable combination of the maximum-entropy (or local
\textit{max-ent}) family shape functions, and the proposal of
a explicit predictor-corrector scheme. We illustrate the advantages of
this approach thorough to simple but challenging test cases, like the
propagation of shock waves in a elastic bar or the response of a block of soil
gradually loaded with gravitational forces.

The article is organized as follows. Section \ref{sec:Notation}
briefly reviews the notation here employed. Next, Section \ref{sec:derivation-mpm}
is devoted to present briefly the governing equations of the elastic
problem, the variational formulation and the Galerkin procedure. In Section \ref{sec:epc-algor-mpm} an explicit predictor-corrector
time integration scheme for MPM is proposed. Section
\ref{sec:local-max-ent} briefly reviews the local \textit{max-ent}
basis functions here employed. In Section
\ref{sec:Application-linear-elasticity-dynamic-problems} applications
to prove the numerical accuracy of the proposed approach are presented. Finally,
conclusions and future research topics are exposed in Section \ref{sec:conclusions}.

\section{Brief note concerning the notation}
\label{sec:Notation}
In what follows, we will adopt the following convention. All the
physical variables involved in this problem are collected in Table
\ref{tab:notation-table}. Three kind of subscript or superscript are used within paper. The
subscript $\qed_p$ is used to define a particle variable. While the
subscript $\qed_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\qed^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\qed}$ and
$\ddot{\qed}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $(\cdot)$ and $(\colon)$ means the first
and second contraction of a tensor, $\Div{\qed}$ denotes the
divergence operator, and finally $\Grad{\qed}$ and $\GradS{\qed}$
denotes the gradient and its symmetric part. Einstain subscripts
convention is adopted therefore repeated index means addition.

\begin{table}
  \centering
  \caption{Physical variables involved in the problem}
  \label{tab:notation-table}
  \begin{tabular}{c c c}
    $\rho$ & Density field & Scalar \\
    $\vect{a}$ & Acceleration field & First order tensor \\
    $\vect{v}$ & Velocity field & First order tensor \\
    $\vect{u}$ & Displacement field & First order tensor \\
    $\vect{x}$ & Global coordinates & First order tensor \\
    $\vect{\xi}$ & Local coordinates & First order tensor \\
    $\tens{\sigma}$ & Cauchy stress tensor & Second order tensor \\
    $\tens{\varepsilon}$ & Cauchy strain tensor & Second order tensor \\
    $\tens{D}$ & Constitutive tensor & Fourth order tensor\\
  \end{tabular}
\end{table}

\section{Derivation of MPM procedure}
\label{sec:derivation-mpm}
The aim of this section is to provide an overview of the standard
explicit MPM algorithm~\cite{Sulsky1994}. In consequence it
is structured as follows: the governing equations will be introduced
in \ref{sec:The boundary-value-problem}, later the variational statement of
the problem will be presented \ref{sec:variational-formulation}, and
finally a discretization procedure thorough Galerkin is performed in
\ref{sec:Galerkin-procedure}. The method has three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation thorough a FEM-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence MPM can be regarded as a
Lagrangian-Eulerian method where particles carries on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation.

\subsection{The boundary-value problem}
\label{sec:The boundary-value-problem}
In MPM the continuum mechanics approach is considered. So on, let
define a continuum domain $\varOmega$ occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and $\partial
\varOmega$ the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field $\vect{u}$ allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field $\vect{u}$ through the following set of governing equations and
restrictions that must be satisfied. First (i) the \textit{compatibility
  equation} that extracts from $\vect{u}$ the strain field is,
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type
$\varGamma_d$. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. The corresponding
conjugate variable for the strain field if the stress field
$\tens{\sigma}$, which satisfies (ii) the \textit{conservation of
  momentum equation}
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type
$\varGamma_n$. Next (iii) the constitutive equation as a linear
application from $\Re^n$ to  $\Re^n$, which relates the
strain tensor with the stress tensor,
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

\subsection{Variational formulation}
\label{sec:variational-formulation}
To write the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vect{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vect{u}^{\psi} \in
  \mathcal{H}^1 \mid \vect{u}^{\psi} = \vect{0}\ \text{on}\ \Gamma_d \}.
\end{equation}
And which satisfies that the Cauchy sequences are convergent in $\varOmega$
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vect{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vect{u} \in
\mathcal{H}^1_0$ such that, for $\vect{u}^{\psi} \in
\mathcal{H}^1_0$, the following holds:
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}},
\end{equation}\\
therefore (\ref{eq:BalanceMomentum_wf}) together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative} represents the weak form
formulation of the problem.

\subsection{Galerkin procedure}
\label{sec:Galerkin-procedure}
In order to arrive to a finite set of equations, in contrast with the
FEM, in MPM a double discretization procedure is performed as we will
describe here below. First, the continuum domain
$\Omega$ is discretized with a finite sum of material points (in the
following particles), each one
represent a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point $\vec{x}_p$ is defined at the centroid
of each $\Omega_p$, figure~\ref{fig:MPM-discretization}.
Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$, but also the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where we recover the Riemann
integral definition as an addition of a finite set of points, and
their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
